### 2018-03-02 submission/manuscript.tex
final re-read edits

Although we expected that the full OTU models would perform the best at
classifying individuals with and without carcinomas, our observations
suggest that both the full and significant OR taxa models performed
equally well. These results suggest that lower level classification to
species and strain may not add extra useful information with respect to
prediction models. This has been suggested in previous literature where
metagenomics did not perform better than 16S rRNA gene sequencing data
at classifying individuals with normal colons and those with carcinomas
(30). One possible reason as to why lower level classification may not
result in better models is that the communities are patchy and higher
level taxonomic information pools some of this patchiness allowing for
better prediction models. There may also be a fair bit of data
redundancy within models that utilize more of the community. An example
of this redundancy would be when we trained models on one study and
tested it on the other studies and the AUCs of the models created with
the select OR taxa performed as well as full taxa models {[}Figure
6B{]}.

Our observations also suggest that a small collection of taxa can
classify diease as well as full OTU-based models but that these taxa
individually perform quite poorly {[}Figure 3{]}. This result supports
the contention that there might be redundancy of function even amongst
the taxa included in the significant OR models. As an example multiple
different microbes could be similarily stimulating the activation of
inflammatory pathways and by doing so exacerbate disease progression.
Multiple reports within the literature have found that different
bacteria, such as \emph{Escherichia coli} and \emph{Fusobacterium
nucleatum}, can worsen similarily worsen inflammation in mouse models of
tumorigenesis (5, 6, 21). Although the inflammatory taxa were patchy in
their importance and presence across studies those that were not
typically associated with inflammation were consistently important for
every study {[}Figure 6{]}. The loss of these taxa (\emph{Ruminococcus}
and \emph{Clostridium XI} in stool and \emph{Dorea} and \emph{Blautia}
in unmatched tissue) is particularly interesting because many are
commonly thought to be beneficial due to their involvement in production
of short chain fatty acids (31--33).

The adenoma models as a whole performed poorly in classifying
individuals with and without adenomas. This outcome is not inconsistent
with what has been published previously (27, 34). However, the modeling
results are at odds with results obtained in Baxter, \emph{et al.} (12).
There are some major differences between the models generated in this
meta-analysis and what was used in this previous report. First, the
prevoius report's models investigated the classification of lesions
(individuals with adenoma or carcinoma) and not adenoma alone. The
Baxter, \emph{et al.} models also contained Fecal Immunoglobulin Test
data while our meta-analysis models only contained 16S rRNA gene
sequencing data. Although being able to classify individuals with
adenomas is important, the classification of advanced adenomas is a more
clinically meaningful diagnostic tool (i.e.~those that are at high risk
of progressing to a carcinoma). It is possible that we might have been
able to detect differences in the bacterial community if advanced
adenomas were separated from adenomas but that data was not available
for the majority of studies analyzed. It is also possible that the
initial changes to the bacterial community are focal to where the
initial adenoma develops and would not be easily assessed with a fecal
sample.

Although stool represents an easy and less invasive way to assess risk,
it is not clear how well this sample reflects adenoma- and carcinoma-
associated microbial communities. Some studies have tried to assess this
in health and disease but are limited by their sample size (17, 35).
Sampling the microbiota directly associated with colon tissue may
provide clearer answers but the colon tissue-based studies did not
provide a clearer understanding of how the microbiota may be associated
with tumors. Generally, the full OTU-based models of unmatched and
matched colon tissue samples were concordant with stool samples showing
that GI resident microbes were the most prevalent in the top 10 most
important variables across study {[}Figure S3{]}. \emph{Fusobacterium}
was not identified consistently across studies and this could be due to
both a small number of studies and a small sample size within these
studies. Additionally, the majority of the colon tissue-based results
were study specific with many of the top 10 taxa being present only in a
single study. The presence of genera associated with contamination (36),
within the top 10 most important variables for the genera and OTU models
is worrying (e.g. \emph{Novosphingobium}, \emph{Acidobacteria Gp2},
\emph{Sphingomonas}, etc.). The low bacterial biomass of tissue samples
coupled with potential contamination and small sample sizes could
explain why these results seem to be more sporadic than the stool
results.

One important caveat to this study is that even though genera associated
with certain species such as \emph{Bacteroides fragilis} and
\emph{Streptococcus gallolyticus} subsp. \emph{gallolyticus} were not
identified, it does not necessarily mean that these specific species are
not important in human CRC (22, 24). There are reports that
\emph{Bacteroides fragilis}, positive for the enterotoxigenic gene, are
found at specific locations along the colon but the samples we were able
to use in this meta-analysis could not identify these types of
differences (37). Additionally, since we are limited in our aggregation
of the data to the genus level, it is not possible to clearly delineate
which species are contributing to overall disease progression. Our
observations are not inconsistent with the previous literature on either
\emph{Bacteroides fragilis} or \emph{Streptococcus gallolyticus} subsp.
\emph{gallolyticus}. As an example, the stool-based full community
models consistently identified the genus \emph{Bacteroides}, as well as
OTUs that classified as \emph{Bacteroides}, to be important model
components across studies. This suggests that even though
\emph{Bacteroides} may not increase the OR of individuals having an
adenoma or carcinoma and may not vary in relative abundance, like
\emph{Fusobacterium}, it is still important in CRC. Additionally,
\emph{Streptococcus gallolyticus} subsp. \emph{gallolyticus} is a
mouth-associated microbe, and the results from this study suggest that
regardless of sample type, mouth-associated genera are commonly
associated with an increased OR for individuals to have a carcinoma
tumor.

Despite these limitations the findings that we present here would not be
possible without performing a meta-analysis. These types of studies can
be a useful tool in microbiota research because they can both validate
existing research and make new discoveries by pooling many independent
investigations together. Yet, it is still difficult to perform these
studies because of inaccessible 16S sequencing data, missing or vague
metadata (e.g.~which samples are carcioma and which are not), varying
sequence quality, and multiple small data sets. Better attention to
these specific problems could help to increase the reproducibility and
replicability of microbiota studies and make it easier to perform these
crucial meta-analyses. Moving forward, meta-analyses will be important
tools to help aggregate and find commonalities across studies when
investigating the microbiota in the context of a specific disease and
more are needed (28, 38--40).

By aggregating together a large collection of studies analyzing both
fecal and colon tissue samples, we are able to provide evidence
supporting the importance of the bacterial community in carcinoma
tumors. Although further validation of the biomarkers presented here
need to be undertaken, the replciability of the AUC of a specific
collection of taxa across multiple studies suggests a strong potential
for the use of the microbiota as a risk stratification tool for
individuals with carcinomas.

@@

\textbf{\emph{Obtaining Data Sets:}} The studies used for this
meta-analysis were identified through the review articles written by
Keku, \emph{et al.} and Vogtmann, \emph{et al.} (41, 42) and additional
studies not mentioned in the reviews were obtained based on the authors'
knowledge of the literature. Studies that used tissue or feces as their
sample source for 454 or Illumina 16S rRNA gene sequencing analysis and
had data sets with sequences available for analysis were included. Some
studies were excluded because they did not have publicly available
sequences or did not have metadata in which the authors were able to
share. After these filtering steps, the following studies remained: Ahn,
\emph{et al.} (11), Baxter, \emph{et al.} (12), Brim, \emph{et al.}
(29), Burns, \emph{et al.} (15), Chen, \emph{et al.} (13), Dejea,
\emph{et al.} (20), Flemer, \emph{et al.} (17), Geng, \emph{et al.}
(19), Hale, \emph{et al.} (27), Kostic, \emph{et al.} (43), Lu, \emph{et
al.} (26), Sanapareddy, \emph{et al.} (25), Wang, \emph{et al.} (14),
Weir, \emph{et al.} (23), and Zeller, \emph{et al.} (16). The Zackular
(44) study was not included because the 90 individuals analyzed within
the study are contained within the larger Baxter study (12). After
sequence processing, all the case samples for the Kostic study had 100
or less sequences remaining and was excluded, leaving a total of 14
studies that analysis could be completed on.

@@

\textbf{\emph{Sequence Processing:}} For the majority of studies, raw
sequences were downloaded from the Sequence Read Archive (SRA)
(\url{ftp://ftp-trace.ncbi.nih.gov/sra/sra-instant/reads/ByStudy/sra/SRP/})
and metadata were obtained by searching the respective accession number
of the study at the following website:
\url{http://www.ncbi.nlm.nih.gov/Traces/study/}. Of the studies that did
not have sequences and metadata on the SRA, data was obtained from DBGap
(n = 1, (11)) and directly from the authors (n = 4, (17, 23, 25, 27)).
Each study was processed using the mothur (v1.39.3) software program
(45) and quality filtering utilized the default methods for both 454 and
Illumina based sequencing. If it was not possible to use the defaults,
the stated quality cut-offs, from the study itself, were used instead.
Sequences that were made up of an artificial combination of two or more
different sequences and commonly known as chimeras were identified and
removed using VSEARCH (46) before \emph{de novo} OTU clustering at 97\%
similarity was completed using the OptiClust algorithm (47).

\textbf{\emph{Study Analysis Overview:}} OTU richness, evenness, and
Shannon diversity were first assessed for differences between controls,
adenoma tumors, and carcinoma turmors using both linear mixed-effect
models and ORs. For each individual study the Bray-Curtis index was used
to assess differences between control-adenoma and control-carcinoma
individuals. Next, all common genera were assessed for differences in
ORs for individuals having an adenoma or carcinoma and corrected for
multiple comparisons using the Benjamini-Hochberg method (48). We then
built Random Forest models based on all genera, all OTUs, or significant
OR taxa (only using taxa still significant after multiple comparison
correction). For both the full genera and significant OR taxa, models
were trained on one study then tested on the remaining studies using
genera-based relative abundances. The OTU-based models were built using
OTU level data and a 10-fold CV over 100 different iterations, based on
random 80/20 splitting of the data, was used to generate a range of
expected AUCs. This process was repeated for every study in the
meta-analysis. Comparisons of the initial trained model AUCs for the
full genera and significant OR taxa were made to the mean AUC generated
from the 100 different 10-fold CV runs of the respective OTU-based
model. For comparisons in which only control versus adenoma individuals
were made, the carcinoma individuals were excluded from each respective
study. Similarly, for comparisons in which control versus carcinoma
individuals were made the adenoma individuals were excluded from each
respective study. For all analysis completed fecal and tissue samples
were kept separate. Within the tissue groups the data were further
divided between samples from the same individual (matched) and those
from different individuals (unmatched).

@@

\textbf{\emph{Statistical Analysis:}} All statistical analysis after
sequence processing utilized the R (v3.4.3) software package (49). For
OTU richness, evenness, and Shannon diversity analysis, values were
power transformed using the rcompanion (v1.11.1) package (50) and then
Z-score normalized using the car (v2.1.6) package (51). Testing for OTU
richness, evenness, and Shannon diversity differences utilized linear
mixed-effect models created using the lme4 (v1.1.15) package (52) to
correct for study, repeat sampling of individuals (tissue only), and 16S
hyper-variable region used. Odds ratios (OR) were analyzed using both
the epiR (v0.9.93) and metafor (v2.0.0) packages (53, 54) by assessing
how many individuals with and without disease were above and below the
overall median value within each specific study. OR significance testing
utilized the chi-squared test. Diversity differences measured by the
Bray-Curtis index utilized the creation of distance matrix and testing
with PERMANOVA executed with the vegan (v2.4.5) package (55). Random
Forest models were built using both the caret (v6.0.78) and randomForest
(v4.6.12) packages (56, 57). All figures were created using both ggplot2
(v2.2.1) and gridExtra (v2.3) packages (58, 59).

@@

\hypertarget{ref-Flynn_preprint_2017}{}
35. \textbf{Flynn, K. J.}, \textbf{M. T. Ruffin}, \textbf{D. K.
Turgeon}, and \textbf{P. D. Schloss}. 2017. Spatial variation of the
native colon microbiota in healthy adults. Cold Spring Harbor
Laboratory.

\hypertarget{ref-Salter_contamination_2014}{}
36. \textbf{Salter, S. J.}, \textbf{M. J. Cox}, \textbf{E. M. Turek},
\textbf{S. T. Calus}, \textbf{W. O. Cookson}, \textbf{M. F. Moffatt},
\textbf{P. Turner}, \textbf{J. Parkhill}, \textbf{N. J. Loman}, and
\textbf{A. W. Walker}. 2014. Reagent and laboratory contamination can
critically impact sequence-based microbiome analyses. BMC Biology.
Springer Nature \textbf{12}.

\hypertarget{ref-Purcell2017}{}
37. \textbf{Purcell, R. V.}, \textbf{J. Pearson}, \textbf{A. Aitchison},
\textbf{L. Dixon}, \textbf{F. A. Frizelle}, and \textbf{J. I. Keenan}.
2017. Colonization with enterotoxigenic bacteroides fragilis is
associated with early-stage colorectal neoplasia. PLOS ONE. Public
Library of Science (PLoS) \textbf{12}:e0171602.

\hypertarget{ref-Sze2016}{}
38. \textbf{Sze, M. A.}, and \textbf{P. D. Schloss}. 2016. Looking for a
signal in the noise: Revisiting obesity and the microbiome. mBio.
American Society for Microbiology \textbf{7}:e01018--16.

\hypertarget{ref-Walters2014}{}
39. \textbf{Walters, W. A.}, \textbf{Z. Xu}, and \textbf{R. Knight}.
2014. Meta-analyses of human gut microbes associated with obesity and
IBD. FEBS Letters. Wiley-Blackwell \textbf{588}:4223--4233.

\hypertarget{ref-Finucane2014}{}
40. \textbf{Finucane, M. M.}, \textbf{T. J. Sharpton}, \textbf{T. J.
Laurent}, and \textbf{K. S. Pollard}. 2014. A taxonomic signature of
obesity in the microbiome? Getting to the guts of the matter. PLoS ONE.
Public Library of Science (PLoS) \textbf{9}:e84689.

\hypertarget{ref-keku_gastrointestinal_2015}{}
41. \textbf{Keku, T. O.}, \textbf{S. Dulal}, \textbf{A. Deveaux},
\textbf{B. Jovov}, and \textbf{X. Han}. 2015. The gastrointestinal
microbiota and colorectal cancer. American Journal of Physiology -
Gastrointestinal and Liver Physiology \textbf{308}:G351--G363.

\hypertarget{ref-vogtmann_epidemiologic_2016}{}
42. \textbf{Vogtmann, E.}, and \textbf{J. J. Goedert}. 2016.
Epidemiologic studies of the human microbiome and cancer. British
Journal of Cancer \textbf{114}:237--242.

\hypertarget{ref-kostic_genomic_2012}{}
43. \textbf{Kostic, A. D.}, \textbf{D. Gevers}, \textbf{C. S.
Pedamallu}, \textbf{M. Michaud}, \textbf{F. Duke}, \textbf{A. M. Earl},
\textbf{A. I. Ojesina}, \textbf{J. Jung}, \textbf{A. J. Bass},
\textbf{J. Tabernero}, \textbf{J. Baselga}, \textbf{C. Liu}, \textbf{R.
A. Shivdasani}, \textbf{S. Ogino}, \textbf{B. W. Birren}, \textbf{C.
Huttenhower}, \textbf{W. S. Garrett}, and \textbf{M. Meyerson}. 2012.
Genomic analysis identifies association of Fusobacterium with colorectal
carcinoma. Genome Research \textbf{22}:292--298.

\hypertarget{ref-zackular_human_2014}{}
44. \textbf{Zackular, J. P.}, \textbf{M. A. M. Rogers}, \textbf{M. T.
Ruffin}, and \textbf{P. D. Schloss}. 2014. The human gut microbiome as a
screening tool for colorectal cancer. Cancer Prevention Research
(Philadelphia, Pa.) \textbf{7}:1112--1121.

\hypertarget{ref-schloss_introducing_2009}{}
45. \textbf{Schloss, P. D.}, \textbf{S. L. Westcott}, \textbf{T.
Ryabin}, \textbf{J. R. Hall}, \textbf{M. Hartmann}, \textbf{E. B.
Hollister}, \textbf{R. A. Lesniewski}, \textbf{B. B. Oakley}, \textbf{D.
H. Parks}, \textbf{C. J. Robinson}, \textbf{J. W. Sahl}, \textbf{B.
Stres}, \textbf{G. G. Thallinger}, \textbf{D. J. Van Horn}, and
\textbf{C. F. Weber}. 2009. Introducing mothur: Open-Source,
Platform-Independent, Community-Supported Software for Describing and
Comparing Microbial Communities. Appl.Environ.Microbiol.
\textbf{75}:7537--7541.

\hypertarget{ref-rognes_vsearch_2016}{}
46. \textbf{Rognes, T.}, \textbf{T. Flouri}, \textbf{B. Nichols},
\textbf{C. Quince}, and \textbf{F. Mahé}. 2016. VSEARCH: A versatile
open source tool for metagenomics. PeerJ \textbf{4}:e2584.

\hypertarget{ref-westcott_opticlust_2017}{}
47. \textbf{Westcott, S. L.}, and \textbf{P. D. Schloss}. 2017.
OptiClust, an Improved Method for Assigning Amplicon-Based Sequence Data
to Operational Taxonomic Units. mSphere \textbf{2}.

\hypertarget{ref-benjamini_controlling_1995}{}
48. \textbf{Benjamini, Y.}, and \textbf{Y. Hochberg}. 1995. Controlling
the false discovery rate: A practical and powerful approach to multiple
testing. Journal of the Royal Statistical Society. Series B
(Methodological) \textbf{57}:289--300.

\hypertarget{ref-r_citation_2017}{}
49. \textbf{R Core Team}. 2017. R: A language and environment for
statistical computing. R Foundation for Statistical Computing, Vienna,
Austria.

\hypertarget{ref-rcompanion_citation_2017}{}
50. \textbf{Mangiafico, S.} 2017. Rcompanion: Functions to support
extension education program evaluation.

\hypertarget{ref-car_citation_2011}{}
51. \textbf{Fox, J.}, and \textbf{S. Weisberg}. 2011. An R companion to
applied regressionSecond. Sage, Thousand Oaks CA.

\hypertarget{ref-lme4_citation_2015}{}
52. \textbf{Bates, D.}, \textbf{M. Mächler}, \textbf{B. Bolker}, and
\textbf{S. Walker}. 2015. Fitting linear mixed-effects models using
lme4. Journal of Statistical Software \textbf{67}:1--48.

>>>>

Although we expected that the full OTU models would perform the best at
classifying individuals with and without carcinomas, our observations
suggest that both the full and significant OR taxa models performed
equally well. These results suggest that lower level classification to
species and strain may not add extra useful information with respect to
prediction models. This has been suggested in previous literature where
metagenomics did not perform better than 16S rRNA gene sequencing data
at classifying individuals with normal colons and those with carcinomas
(30). One possible reason as to why lower level classification may not
result in better models is that the communities are patchy and higher
level taxonomic information pools some of this patchiness, allowing for
better prediction models. There may also be a fair bit of data
redundancy within models that utilize more of the community. An example
of this redundancy would be when we trained models on one study and
tested it on the other studies and the AUCs of the models created with
the select OR taxa performed as well as full taxa models {[}Figure
6B{]}.

Our observations also suggest that a small collection of taxa can
classify disease as well as full OTU-based models but that these taxa
individually perform quite poorly {[}Figure 3{]}. This result supports
the contention that there might be redundancy of function amongst the
taxa included in the significant OR models. As an example, multiple
different microbes could be similarily stimulating the activation of
inflammatory pathways and by doing so exacerbate disease progression.
Multiple reports within the literature have found that different
bacteria, such as \emph{Escherichia coli} and \emph{Fusobacterium
nucleatum}, can similarily worsen inflammation in mouse models of
tumorigenesis (5, 6, 21). Although the inflammatory taxa were patchy in
their importance and presence across studies those that were not
typically associated with inflammation were consistently important for
every study {[}Figure 6{]}. The loss of these taxa (\emph{Ruminococcus}
and \emph{Clostridium XI} in stool and \emph{Dorea} and \emph{Blautia}
in unmatched tissue) is particularly interesting because many are
commonly thought to be beneficial due to their involvement in production
of short chain fatty acids (31--33).

The adenoma models as a whole performed poorly in classifying
individuals with and without adenomas. This outcome is not inconsistent
with what has been published previously (27, 34). However, the modeling
results are at odds with results obtained in Baxter, \emph{et al.} (12).
There are some major differences between the models generated in this
meta-analysis and what was used in Baxter, \emph{et al}. First, the
prevoius report's models investigated the classification of lesions
(individuals with adenoma or carcinoma) and not adenoma alone. The
Baxter, \emph{et al.} models also contained Fecal Immunoglobulin Test
data while our meta-analysis models only contained 16S rRNA gene
sequencing data. Although being able to classify individuals with
adenomas is important, the classification of advanced adenomas is a more
clinically meaningful diagnostic tool (i.e.~those that are at highest
risk of progressing to a carcinoma). It is possible that we might have
been able to detect differences in the bacterial community if advanced
adenomas were separated from adenomas but that data was not available
for the majority of studies analyzed. Additionally, the initial changes
to the bacterial community could be focal to where the initial adenoma
develops and would not be easily assessed with a fecal sample.

Although stool represents an easy and less invasive way to assess risk,
it is not clear how well this sample reflects adenoma- and carcinoma-
tissue associated microbial communities. The colon tissue-based studies
did not provide a clearer understanding of how the microbiota may be
associated with tumors. Generally, the full OTU-based models of
unmatched and matched colon tissue samples were concordant with stool
samples showing that GI resident microbes were the most prevalent in the
top 10 most important variables across study {[}Figure S3{]}.
\emph{Fusobacterium} was not consistently identified across studies and
this could be due to both a small number of studies and a small sample
size within these studies. The majority of the colon tissue-based
results were also study specific with many of the top 10 taxa being
present only in a single study. Additionally, the presence of genera
associated with contamination, within the top 10 most important
variables for the genera and OTU models is worrying (e.g.
\emph{Novosphingobium}, \emph{Acidobacteria Gp2}, \emph{Sphingomonas},
etc. (35)). The low bacterial biomass of tissue samples coupled with
potential contamination and small sample sizes could explain why these
results seem to be more sporadic than the stool results.

One important caveat to this study is that even though genera associated
with certain species such as \emph{Bacteroides fragilis} and
\emph{Streptococcus gallolyticus} subsp. \emph{gallolyticus} were not
identified, it does not necessarily mean that these specific species are
not important in human CRC (22, 24). There are reports that
\emph{Bacteroides fragilis}, positive for the enterotoxigenic gene, are
found at specific locations along the colon but the samples we were able
to use in this meta-analysis could not identify these types of
differences (36). Additionally, since we are limited in our aggregation
of the data to the genus level, it is not possible to clearly delineate
which species are contributing to overall disease progression. Our
observations are not inconsistent with the previous literature on either
\emph{Bacteroides fragilis} or \emph{Streptococcus gallolyticus} subsp.
\emph{gallolyticus}. As an example, the stool-based full community
models consistently identified the \emph{Bacteroides} taxa to be
important model components across studies. This suggests that even
though \emph{Bacteroides} may not increase the OR of individuals having
an adenoma or carcinoma and may not vary in relative abundance, like
\emph{Fusobacterium}, it is still important in CRC.

Despite these limitations to the findings, meta-analyses are a useful
tool in microbiota research because they can both validate existing
research and make new discoveries by pooling many independent
investigations together. Yet, it is still difficult to perform these
studies because of inaccessible 16S sequencing data, missing or vague
metadata (e.g.~which samples are carcioma and which are not), varying
sequence quality, and multiple small data sets. Better attention to
these specific problems could help to increase the reproducibility and
replicability of microbiota studies and make it easier to perform these
crucial meta-analyses. Moving forward, meta-analyses will be important
tools to help aggregate and find commonalities across studies when
investigating the microbiota in the context of a specific disease and
more are needed (28, 37--39).

By aggregating together a large collection of studies analyzing both
fecal and colon tissue samples, we are able to provide evidence
supporting the importance of the bacterial community in carcinoma
tumors. Although further validation of the biomarkers presented here
need to be undertaken, the replicability of the AUC of a specific
collection of taxa across multiple studies suggests a strong potential
for the use of the microbiota as a risk stratification tool for
individuals with carcinomas.

@@

\textbf{\emph{Obtaining Data Sets:}} The studies used for this
meta-analysis were identified through the review articles written by
Keku, \emph{et al.} and Vogtmann, \emph{et al.} (40, 41) and additional
studies not mentioned in the reviews were obtained based on the authors'
knowledge of the literature. Studies that used tissue or feces as their
sample source for 454 or Illumina 16S rRNA gene sequencing analysis and
had data sets with sequences available for analysis were included. Some
studies were excluded because they did not have publicly available
sequences or did not have metadata in which the authors were able to
share. After these filtering steps, the following studies remained: Ahn,
\emph{et al.} (11), Baxter, \emph{et al.} (12), Brim, \emph{et al.}
(29), Burns, \emph{et al.} (15), Chen, \emph{et al.} (13), Dejea,
\emph{et al.} (20), Flemer, \emph{et al.} (17), Geng, \emph{et al.}
(19), Hale, \emph{et al.} (27), Kostic, \emph{et al.} (42), Lu, \emph{et
al.} (26), Sanapareddy, \emph{et al.} (25), Wang, \emph{et al.} (14),
Weir, \emph{et al.} (23), and Zeller, \emph{et al.} (16). The Zackular
(43) study was not included because the 90 individuals analyzed within
the study are contained within the larger Baxter study (12). After
sequence processing, all the case samples for the Kostic study had 100
or less sequences remaining and was excluded, leaving a total of 14
studies that analysis could be completed on.

@@

\textbf{\emph{Sequence Processing:}} For the majority of studies, raw
sequences were downloaded from the Sequence Read Archive (SRA)
(\url{ftp://ftp-trace.ncbi.nih.gov/sra/sra-instant/reads/ByStudy/sra/SRP/})
and metadata were obtained by searching the respective accession number
of the study at the following website:
\url{http://www.ncbi.nlm.nih.gov/Traces/study/}. Of the studies that did
not have sequences and metadata on the SRA, data was obtained from DBGap
(n = 1, (11)) and directly from the authors (n = 4, (17, 23, 25, 27)).
Each study was processed using the mothur (v1.39.3) software program
(44) and quality filtering utilized the default methods for both 454 and
Illumina based sequencing. If it was not possible to use the defaults,
the stated quality cut-offs, from the study itself, were used instead.
Sequences that were made up of an artificial combination of two or more
different sequences and commonly known as chimeras were identified and
removed using VSEARCH (45) before \emph{de novo} OTU clustering at 97\%
similarity was completed using the OptiClust algorithm (46).

\textbf{\emph{Study Analysis Overview:}} OTU richness, evenness, and
Shannon diversity were first assessed for differences between controls,
adenoma tumors, and carcinoma turmors using both linear mixed-effect
models and ORs. For each individual study the Bray-Curtis index was used
to assess differences between control-adenoma and control-carcinoma
individuals. Next, all common genera were assessed for differences in
ORs for individuals having an adenoma or carcinoma and corrected for
multiple comparisons using the Benjamini-Hochberg method (47). We then
built Random Forest models based on all genera, all OTUs, or significant
OR taxa (only using taxa still significant after multiple comparison
correction). For both the full genera and significant OR taxa, models
were trained on one study then tested on the remaining studies using
genera-based relative abundances. The OTU-based models were built using
OTU level data and a 10-fold CV over 100 different iterations, based on
random 80/20 splitting of the data, was used to generate a range of
expected AUCs. This process was repeated for every study in the
meta-analysis. Comparisons of the initial trained model AUCs for the
full genera and significant OR taxa were made to the mean AUC generated
from the 100 different 10-fold CV runs of the respective OTU-based
model. For comparisons in which only control versus adenoma individuals
were made, the carcinoma individuals were excluded from each respective
study. Similarly, for comparisons in which control versus carcinoma
individuals were made the adenoma individuals were excluded from each
respective study. For all analysis completed fecal and tissue samples
were kept separate. Within the tissue groups the data were further
divided between samples from the same individual (matched) and those
from different individuals (unmatched).

@@

\textbf{\emph{Statistical Analysis:}} All statistical analysis after
sequence processing utilized the R (v3.4.3) software package (48). For
OTU richness, evenness, and Shannon diversity analysis, values were
power transformed using the rcompanion (v1.11.1) package (49) and then
Z-score normalized using the car (v2.1.6) package (50). Testing for OTU
richness, evenness, and Shannon diversity differences utilized linear
mixed-effect models created using the lme4 (v1.1.15) package (51) to
correct for study, repeat sampling of individuals (tissue only), and 16S
hyper-variable region used. Odds ratios (OR) were analyzed using both
the epiR (v0.9.93) and metafor (v2.0.0) packages (52, 53) by assessing
how many individuals with and without disease were above and below the
overall median value within each specific study. OR significance testing
utilized the chi-squared test. Diversity differences measured by the
Bray-Curtis index utilized the creation of distance matrix and testing
with PERMANOVA executed with the vegan (v2.4.5) package (54). Random
Forest models were built using both the caret (v6.0.78) and randomForest
(v4.6.12) packages (55, 56). All figures were created using both ggplot2
(v2.2.1) and gridExtra (v2.3) packages (57, 58).

@@

\hypertarget{ref-Salter_contamination_2014}{}
35. \textbf{Salter, S. J.}, \textbf{M. J. Cox}, \textbf{E. M. Turek},
\textbf{S. T. Calus}, \textbf{W. O. Cookson}, \textbf{M. F. Moffatt},
\textbf{P. Turner}, \textbf{J. Parkhill}, \textbf{N. J. Loman}, and
\textbf{A. W. Walker}. 2014. Reagent and laboratory contamination can
critically impact sequence-based microbiome analyses. BMC Biology.
Springer Nature \textbf{12}.

\hypertarget{ref-Purcell2017}{}
36. \textbf{Purcell, R. V.}, \textbf{J. Pearson}, \textbf{A. Aitchison},
\textbf{L. Dixon}, \textbf{F. A. Frizelle}, and \textbf{J. I. Keenan}.
2017. Colonization with enterotoxigenic bacteroides fragilis is
associated with early-stage colorectal neoplasia. PLOS ONE. Public
Library of Science (PLoS) \textbf{12}:e0171602.

\hypertarget{ref-Sze2016}{}
37. \textbf{Sze, M. A.}, and \textbf{P. D. Schloss}. 2016. Looking for a
signal in the noise: Revisiting obesity and the microbiome. mBio.
American Society for Microbiology \textbf{7}:e01018--16.

\hypertarget{ref-Walters2014}{}
38. \textbf{Walters, W. A.}, \textbf{Z. Xu}, and \textbf{R. Knight}.
2014. Meta-analyses of human gut microbes associated with obesity and
IBD. FEBS Letters. Wiley-Blackwell \textbf{588}:4223--4233.

\hypertarget{ref-Finucane2014}{}
39. \textbf{Finucane, M. M.}, \textbf{T. J. Sharpton}, \textbf{T. J.
Laurent}, and \textbf{K. S. Pollard}. 2014. A taxonomic signature of
obesity in the microbiome? Getting to the guts of the matter. PLoS ONE.
Public Library of Science (PLoS) \textbf{9}:e84689.

\hypertarget{ref-keku_gastrointestinal_2015}{}
40. \textbf{Keku, T. O.}, \textbf{S. Dulal}, \textbf{A. Deveaux},
\textbf{B. Jovov}, and \textbf{X. Han}. 2015. The gastrointestinal
microbiota and colorectal cancer. American Journal of Physiology -
Gastrointestinal and Liver Physiology \textbf{308}:G351--G363.

\hypertarget{ref-vogtmann_epidemiologic_2016}{}
41. \textbf{Vogtmann, E.}, and \textbf{J. J. Goedert}. 2016.
Epidemiologic studies of the human microbiome and cancer. British
Journal of Cancer \textbf{114}:237--242.

\hypertarget{ref-kostic_genomic_2012}{}
42. \textbf{Kostic, A. D.}, \textbf{D. Gevers}, \textbf{C. S.
Pedamallu}, \textbf{M. Michaud}, \textbf{F. Duke}, \textbf{A. M. Earl},
\textbf{A. I. Ojesina}, \textbf{J. Jung}, \textbf{A. J. Bass},
\textbf{J. Tabernero}, \textbf{J. Baselga}, \textbf{C. Liu}, \textbf{R.
A. Shivdasani}, \textbf{S. Ogino}, \textbf{B. W. Birren}, \textbf{C.
Huttenhower}, \textbf{W. S. Garrett}, and \textbf{M. Meyerson}. 2012.
Genomic analysis identifies association of Fusobacterium with colorectal
carcinoma. Genome Research \textbf{22}:292--298.

\hypertarget{ref-zackular_human_2014}{}
43. \textbf{Zackular, J. P.}, \textbf{M. A. M. Rogers}, \textbf{M. T.
Ruffin}, and \textbf{P. D. Schloss}. 2014. The human gut microbiome as a
screening tool for colorectal cancer. Cancer Prevention Research
(Philadelphia, Pa.) \textbf{7}:1112--1121.

\hypertarget{ref-schloss_introducing_2009}{}
44. \textbf{Schloss, P. D.}, \textbf{S. L. Westcott}, \textbf{T.
Ryabin}, \textbf{J. R. Hall}, \textbf{M. Hartmann}, \textbf{E. B.
Hollister}, \textbf{R. A. Lesniewski}, \textbf{B. B. Oakley}, \textbf{D.
H. Parks}, \textbf{C. J. Robinson}, \textbf{J. W. Sahl}, \textbf{B.
Stres}, \textbf{G. G. Thallinger}, \textbf{D. J. Van Horn}, and
\textbf{C. F. Weber}. 2009. Introducing mothur: Open-Source,
Platform-Independent, Community-Supported Software for Describing and
Comparing Microbial Communities. Appl.Environ.Microbiol.
\textbf{75}:7537--7541.

\hypertarget{ref-rognes_vsearch_2016}{}
45. \textbf{Rognes, T.}, \textbf{T. Flouri}, \textbf{B. Nichols},
\textbf{C. Quince}, and \textbf{F. Mahé}. 2016. VSEARCH: A versatile
open source tool for metagenomics. PeerJ \textbf{4}:e2584.

\hypertarget{ref-westcott_opticlust_2017}{}
46. \textbf{Westcott, S. L.}, and \textbf{P. D. Schloss}. 2017.
OptiClust, an Improved Method for Assigning Amplicon-Based Sequence Data
to Operational Taxonomic Units. mSphere \textbf{2}.

\hypertarget{ref-benjamini_controlling_1995}{}
47. \textbf{Benjamini, Y.}, and \textbf{Y. Hochberg}. 1995. Controlling
the false discovery rate: A practical and powerful approach to multiple
testing. Journal of the Royal Statistical Society. Series B
(Methodological) \textbf{57}:289--300.

\hypertarget{ref-r_citation_2017}{}
48. \textbf{R Core Team}. 2017. R: A language and environment for
statistical computing. R Foundation for Statistical Computing, Vienna,
Austria.

\hypertarget{ref-rcompanion_citation_2017}{}
49. \textbf{Mangiafico, S.} 2017. Rcompanion: Functions to support
extension education program evaluation.

\hypertarget{ref-car_citation_2011}{}
50. \textbf{Fox, J.}, and \textbf{S. Weisberg}. 2011. An R companion to
applied regressionSecond. Sage, Thousand Oaks CA.

\hypertarget{ref-lme4_citation_2015}{}
51. \textbf{Bates, D.}, \textbf{M. Mächler}, \textbf{B. Bolker}, and
\textbf{S. Walker}. 2015. Fitting linear mixed-effects models using
lme4. Journal of Statistical Software \textbf{67}:1--48.

\hypertarget{ref-epir_citation_2017}{}
52. \textbf{Telmo Nunes, M. S. with contributions from}, \textbf{C.
Heuer}, \textbf{J. Marshall}, \textbf{J. Sanchez}, \textbf{R. Thornton},
\textbf{J. Reiczigel}, \textbf{J. Robison-Cox}, \textbf{P. Sebastiani},
\textbf{P. Solymos}, \textbf{K. Yoshida}, \textbf{G. Jones}, \textbf{S.
Pirikahu}, \textbf{S. Firestone}, and \textbf{R. Kyle.} 2017. EpiR:
Tools for the analysis of epidemiological data.

\hypertarget{ref-metafor_citation_2010}{}
53. \textbf{Viechtbauer, W.} 2010. Conducting meta-analyses in R with
the metafor package. Journal of Statistical Software \textbf{36}:1--48.
